Using cosmic variance to constrain the dark matter halo mass of Lyman-alpha emitting galaxies at $z$=3.1

### 2013-04-19 paper/ms.tex
udpated paper

\title[Dark Matter Halo Mass for LAEs  at $z=3.1$]{Using cosmic
  variance to constrain the
  dark matter halo mass of Lyman-alpha emitting galaxies at $z$=3.1}
\author[~J.~E. Forero-Romero and J. Mejia]{
\parbox[t]{\textwidth}{\raggedright 
  Julian Mej\'ia$^{2}$ and
Jaime E. Forero-Romero$^{1}$}
\vspace*{6pt}\\
$^{1}$\\
$^{2}$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia 
}

@@

\begin{abstract}
We use simulations and recent observations to find the characteristic
mass of dark matter halos hosting Lyman-Alpha Emitting (LAE) galaxies
at a redshift of $z=3.1$. The method is based on matching the
statistics for the number density cosmic variance between observed and
mock fields. We build a model for LAEs on top of a large N-body
cosmological simulation (Bolshoi) where the basic principle is that a
dark matter halo can only host one LAE with a probability $f_{\rm occ}$ if its
mass is found withing a certain range mass range delimited by two
threshold values, $M_{\rm min}$ and $M_{\rm max}$. We generate
hundreds of mock catalogs in a thourough exploration of the parameter
space of this model. These catalogs are constructed in such a way as
to reproduce the spatial correlation between the different observed
fields. We find that the models that best match the observed cosmic
variance statistics are those with halo masses in the range $M_{\rm
  min}=$ and $M_{\rm max}=$ with and occupation fraction that scales
as $f_{\rm occ}=$ with the minimum mass. We find that the angular
correlation function is in agreement with the observational
constraints mostly due to the cosmic variance in the small angular
fields where this statistics has been computed so far. We also present
predictions for future measurements of the correlation function on
larger continous fields. We make available the mock data for the best
models in a public repository.  
\end{abstract}

@@

@@

\cite{Hayes2010} constrained the value of $f_{\rm occ}$ at $z=2.2$ to
be $f_{\rm occ}=0.10$. This estimation was based on blind surveys of
the H$\alpha$ and Lyman alpha line with the European Southern
Observatorio (ESO) Very Large Telescope (VLT). Using corrections by
extinction to obtain an estimate for the intrinsic H$\alpha$
luminosity, and using values for the theoretical expectation of the
ratio Lyman$\alpha$/H$\alpha$ they derive an bulk escape fraction for
the Lyman$\alpha$ radiation of $f_{\rm esc}=(5.3\pm 3.8)\%$ or $f_{\rm
esc}=(10.7\pm 2.8)$ if a different dust correction is used. From this
results they derive that almost $90\%$ of the star forming galaxies
emit insufficient Lyman-alpha to be detected, effectively setting the
occupation fraction to be $f_{\rm occ}=0.10$. Obervations are
consistent with these numbers are
consistent with being constant for every galaxy regardless of its
luminosity.

For the cosmological parameters used in this \documentname the age of
the universe between $z=3.1$ and $z=2.2$ has changed by $\sim 1$ Gyr,
it is reasonable to assume that the physical conditions that determine
the escape fraction $f_{\rm esc}$ and the occupation fraction $f_{\rm
  occ}$ remain constant. This assumption allows us to further pick
models that have an occupation fraction of $f_{\rm occ}=0.10$. Under
this selection only $18$ models can be selected. Considering an
occupation fraction $f_{\rm occ}=0.20$ another $39$ models can be
considered. The minimum and maximum mass are found in the appendix in
Table XX.

>>>>

\title[Dark Matter Halo Mass for LAEs  at $z=3.1$]{Using cosmic
  variance to constrain the dark matter halo mass of Lyman-alpha emitting galaxies at $z$=3.1}

\author[~J.~E. Forero-Romero and J. Mejia]{
\parbox[t]{\textwidth}{\raggedright 
  Julian Mej\'ia$^{2}$ and
Jaime E. Forero-Romero$^{1}$}
\vspace*{6pt}\\
$^{1}$\\
$^{2}$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia 
}

@@

\begin{abstract}
We use cosmological N-body simulations to find the characteristic
mass of dark matter halos hosting Lyman-Alpha Emitting (LAE) galaxies at a redshift of $z=3.1$. The method is based on matching the statistics for the number density between mock and observed fields. The mock catalogs are constructed using a simple model where the basic principle is that a dark matter halo can only host one LAE with a probability $f_{\rm occ}$ if its
mass is found withing a certain range mass range delimited by two
threshold values, $M_{\rm min}$ and $M_{\rm max}$. We generate
hundreds of mock catalogs in a thourough exploration of the parameter space of this model. These catalogs are constructed in such a way as o reproduce the observed areas and spatial correlations between the different observed fields. We find that the models that are consistent the observed cosmic variance statistics are those with halo masses in the range $M_{\rm
  min}=$ and $M_{\rm max}=$ with and occupation fraction that scales as $f_{\rm occ}=$ with the minimum mass. We explore three additional constraints to narrow these range. First we select the models with the highest possible number of mocks consistent with observations. Second we use observational constraints on the occupation fraction. Third, we compute the angular correlation function in the mocks against observational values. The first two conditions narrow down the space parameter to $M_{\rm min}=$ and $M_{\rm max}$, $f_{\rm occ}$. Although we find that the angular correlation function is in agreement with the observational constraints mostly due to the cosmic variance in the small angular fields where this statistics has been computed so far, this statistics does not allows us to put tighter constraints on the model parameters. We also present predictions for future measurements of the correlation function on larger continous fields. We make available the mock data for the best
models in a public repository.  
Implications for galaxy formation models?
\end{abstract}

@@

In what follows we will focus our discussion on the mocks constructed with the {\tt{Match}} and {\tt{Random}} methods.

\section{Discussion}

The raw results of our experiments yield a prefered range for
halo masses hosting LAES is in the range $10^{10}\hMsun<M_{\rm
  min}<10^{11.5}\hMsun$, without any limitation in the  maximum mass, with an average occupation fraction as a function of the minimum mass with an approximate dependence of the kind $f_{occ}\approx 0.1 (M_{\rm min}/10^{10})^{XX}$. These results discard models that were nevertheless disfavored from the very beginning based on the results of the mass functions (see Figure 1.).

In this discussion section we consider three different ways to impose tighter constraints by making use of the results we have derived so far and using additional statistical and astrophysical constraints. The first is based on
models with high fractions of mock surveys that are consistent with
observations. The second uses observational results on the occupation
fraction. The third exploits the observational information in the Angular
Correlation Function (ACF).

\subsection{Models with the highest success rates}
\begin{figure*}
\begin{center}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/Fig4_match_P5.pdf}
\hspace{5mm}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/Fig4_random_P5.pdf}
\end{center} 
\caption{ Number of models with a mininium number of mock survey
  realizations that are consistent with observations.
  \label{figure:high_success_rate}.}  
\end{figure*}

For each model there are 15 different mock survey realizations of mock. In
the loose constraints of the previous sections, each model that had at
least one (1) realization of mock survey with $P>0.05$ was considered as consistent with observations.

In Figure XX we show the number of models that have at least $n$ realizations with $P>0.05$ for the {\texttt match} and {\texttt random} methods.  This shows that there are around $550$ to $600$ different models that have at leaset one mock survey realization consistent with observations. At the other extreme, there are $80$ to $100$ models with all the 15 realizations with $P>0.05$. Here we focuse on the latter models. The best models represent $\sim 15\%$ of the number of initially considered good models.

In Figure XX we present the locii of these models in the parameter space $M_{\rm min}-M_{\rm max}$ and $M_{\rm min}-f_{\rm occ}$.

@@

We now impose a different restriction using the observational results by \cite{Hayes2010}. These authors constrained the value of $f_{\rm occ}$ at $z=2.2$ to be $f_{\rm occ}=0.10$. This estimation was based on blind surveys of
the H$\alpha$ and Lyman alpha line with the European Southern
Observatorio (ESO) Very Large Telescope (VLT). Using corrections by
extinction to obtain an estimate for the intrinsic H$\alpha$
luminosity, and using values for the theoretical expectation of the
ratio Lyman$\alpha$/H$\alpha$ they derive an bulk escape fraction for
the Lyman$\alpha$ radiation of $f_{\rm esc}=(5.3\pm 3.8)\%$ or $f_{\rm
esc}=(10.7\pm 2.8)$ if a different dust correction is used. The authors show that the luminosity function for LAEs at $z=2.2$ is consistent with the escape fraction being constant for every galaxy regardless of its luminosity. From this results they derive that almost $90\%$ of the star forming galaxies emit insufficient Lyman-alpha to be detected, effectively setting the occupation fraction to be $f_{\rm occ}=0.10$.

For the cosmological parameters used in this \documentname the age of
the universe between $z=3.1$ and $z=2.2$ has changed by $\sim 1$ Gyr. We assume that the physical conditions that determine
the escape fraction $f_{\rm esc}$ and the occupation fraction $f_{\rm  occ}$ remain constant over that time scale. This assumption allows us to further pick models that have an occupation fraction of $f_{\rm occ}=0.10$. Under
this selection only $18$ models can be selected. Considering an
occupation fraction $f_{\rm occ}=0.20$ another $39$ models can be
considered. This constraint help us to select $\sim 10\%$ of the original models that were considered as consistent with observations in the previous section.

Figure XX shows the prefered models in the planes $M_{\rm min}-M_{\rm max}$ and $M_{\rm min}-f_{\rm occ}$ for the {\tt match} and {\tt random} methods. The list for the model parameters is found in the appendix in Table XX.

In terms of the constraints done in the previous sub-section, we find that the models in this region of parameter space have an average number of XX mock catalogs consistent with observations.
